LaMachine: A meta-distribution for NLP software

### 2019-01-29 docs/lamachine_clarin2018.tex
[paper] fix casing

\begin{abstract}
We introduce LaMachine, a unified Natural Language Processing (NLP) open-source software distribution to facilitate the
installation and deployment of a large amount of software projects that have been developed in the scope of the
CLARIN-NL project and its current successor CLARIAH. Special attention is paid to encouragement of good software
development practices and reuse of established infrastructure in the scientific and open-source software development
community. We explain what Lamachine is, how it can be used and the technical details. We also compare Lamachine to alternative software distributions and discuss its advantages and limitations. We illustrate how LaMachine can be used in two case studies, one in an exploratory text mining project at the Dutch Health Inspectorate where LaMachine was applied to create a research environment for automatic text analysis for health care quality monitoring, and a second case where Lamachine was used to create a workspace for a one-week, intense collaboration by a diverse research team.
\end{abstract}

@@

 This paper provides a detailed description of Lamachine, its purpose and functionality and the technical details. We
 discuss the advantages and limitations of LaMachine compared to alternative software distributions in Section \ref{sec:comparison}. We demonstrate how LaMachine can create a fully functioning and standalone research environment for text mining  and NLP for Dutch texts in two use cases in Section \ref{sec:case}.

 \section{What is Lamachine?}

@@

The nomenclature is admittedly a bit confusing, but the notion of version discussed in this section refers to the
versions of the various software packages inside LaMachine. It should not be
confused with the actual \emph{version release number} of Lamachine as a whole, which is the version number assigned to
the collective of installation scripts LaMachine provides, and which marks LaMachine releases.

@@

LaMachine offered a convenient platform for a range of different explorations and experiments in the area of NLP and
text mining.  However, for some situations LaMachine, or rather Linux in general, was not a good fit for the audience of
the workshop: For team members who did not have experience with a non-Windows environment, LaMachine was not a suitable
or useful tool. The limit of LaMachine was also reached for members who wanted to use desktop text editors with a
graphical user interface as this is not offered by LaMachine. Moreover, we did not manage to get X-forwarding working in the
Ubuntu Linux VM and after a few attempts the team gave up on resolving this issue due to time pressure. This, also
demonstrates that fine-tuning the configuration of certain aspects of LaMachine, but especially beyond LaMachine, is beyond the reach of a data scientist
without system administration skills. This certainly also applies also to the installation as a whole in the
SURFsara context, which involved things like the partitioning, formatting and mounting of (virtual) drives and setting
up user accounts on the shared VM, all of which require some system administration skills and are too context-specific
to be within the scope of LaMachine.
Lamachine was convenient and speeded up writing code as the most common scientific data-related packages are already present in Lamachine.
 %verwerkt <-- <iris> it moet nog wat beter geformuleerd worden maar je snaprt wat ik bedoel
%@iris: hebben jullie verder geen gebruik proberen te maken van de Jupyter Lab omgeving? Dat biedt in ieder geval iets
%grafisch wat via web te benaderen is?
%@maarten helaas lukte dat ook niet, dat liep bij mij steeds vast. - ik heb opmerking van alex over tijdswinst toegevoefd

>>>>

\begin{abstract}
We introduce LaMachine, a unified Natural Language Processing (NLP) open-source software distribution to facilitate the
installation and deployment of a large amount of software projects that have been developed in the scope of the
CLARIN-NL project and its current successor CLARIAH. Special attention is paid to encouragement of good software
development practices and reuse of established infrastructure in the scientific and open-source software development
community. We explain what LaMachine is, how it can be used and the technical details. We also compare LaMachine to alternative software distributions and discuss its advantages and limitations. We illustrate how LaMachine can be used in two case studies, one in an exploratory text mining project at the Dutch Health Inspectorate where LaMachine was applied to create a research environment for automatic text analysis for health care quality monitoring, and a second case where LaMachine was used to create a workspace for a one-week, intense collaboration by a diverse research team.
\end{abstract}

@@

 This paper provides a detailed description of LaMachine, its purpose and functionality and the technical details. We
 discuss the advantages and limitations of LaMachine compared to alternative software distributions in Section \ref{sec:comparison}. We demonstrate how LaMachine can create a fully functioning and standalone research environment for text mining  and NLP for Dutch texts in two use cases in Section \ref{sec:case}.

 \section{What is LaMachine?}

@@

The nomenclature is admittedly a bit confusing, but the notion of version discussed in this section refers to the
versions of the various software packages inside LaMachine. It should not be
confused with the actual \emph{version release number} of LaMachine as a whole, which is the version number assigned to
the collective of installation scripts LaMachine provides, and which marks LaMachine releases.

@@

LaMachine offered a convenient platform for a range of different explorations and experiments in the area of NLP and
text mining.  However, for some situations LaMachine, or rather Linux in general, was not a good fit for the audience of
the workshop: For team members who did not have experience with a non-Windows environment, LaMachine was not a suitable
or useful tool. The limit of LaMachine was also reached for members who wanted to use desktop text editors with a
graphical user interface as this is not offered by LaMachine. Moreover, we did not manage to get X-forwarding working in the
Ubuntu Linux VM and after a few attempts the team gave up on resolving this issue due to time pressure. This, also
demonstrates that fine-tuning the configuration of certain aspects of LaMachine, but especially beyond LaMachine, is beyond the reach of a data scientist
without system administration skills. This certainly also applies also to the installation as a whole in the
SURFsara context, which involved things like the partitioning, formatting and mounting of (virtual) drives and setting
up user accounts on the shared VM, all of which require some system administration skills and are too context-specific
to be within the scope of LaMachine.
LaMachine was convenient and speeded up writing code as the most common scientific data-related packages are already present in LaMachine.
 %verwerkt <-- <iris> it moet nog wat beter geformuleerd worden maar je snaprt wat ik bedoel
%@iris: hebben jullie verder geen gebruik proberen te maken van de Jupyter Lab omgeving? Dat biedt in ieder geval iets
%grafisch wat via web te benaderen is?
%@maarten helaas lukte dat ook niet, dat liep bij mij steeds vast. - ik heb opmerking van alex over tijdswinst toegevoefd
